---
title: "Installation"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

<hr>

To start using OpenRepGrid you need to go through three steps: You need to 

1. install R 
2. install the OpenRepGrid package
3. load the package. 

The video guide at the bottom also shows how to do that.

### Installing R and RStudio

The prerequisite for installing the `OpenRepGrid` package is that R has been installed on your computer. R can be downloaded from the CRAN server for all three big operating systems (Windows, MacOS, Linux). Please follow the installation instructions on the R website. After installing R you can proceed.


### Installing the package

Once R is installed you need to install the `OpenRepGrid` package on your computer. This is a standard step that has to be done for any additional R package you would like to use. There are several options how to do the installing. The simplest is to install the package via the R menu (steps 1 to 3).

1. click install package(s) in the menu
2. select a location near to you
3: select OpenRepGrid

Alternatively you may also use the R command line to install the package. Just type in the following code:
install.packages("OpenRepGrid", dependencies=TRUE)
The installation has to be done once. Afterwards the package is stored on your computer and can be used by R.

### Loading the package

Given that R and the package are installed on your computer you can start using its features. To do this, we need to tell R to load the package. This is a step you have do do **each time** when opening R. As there might be a lot of packages installed on your computer for memory optimization reasons R will only load the most important ones automatically. If you want to use any of the other packages (like e.g. `OpenRepGrid`) you need to tell R to load the additional package to be ready for use.

To load the `OpenRepGrid` package type:

```{r eval=FALSE}
library(OpenRepGrid)
```

That's it, now you can start analyzing your grids.

### Video tutorial

<iframe src="//player.vimeo.com/video/27903566" width="600" height="450" frameborder="0" webkitallowfullscreen mozallowfullscreen allowfullscreen></iframe>

### Development version

The version of the package hosted on CRAN is a checked release. The latest development version is hosted at github. To install the latest version with the features currently under development you need to follow three steps:

1. Make sure you have the latest version of R installed.
2. Make sure the following packages are installed. If not install them by typing
```{r eval=FALSE}
install.packages(c("colorspace", "plyr", "abind", "rgl", "psych", "XML"))
```
3. To download and install the OpenRepGrid development version from github you can use a function from the `devtools` package.
```{r eval=FALSE}
library(devtools)
install_github("OpenRepGrid", "markheckmann") 
```



